---
title: "HW 07"
author: "Matt Scheffel"
date: "2022-10-17"
output:
  pdf_document: default
  html_document: default
---

# Matt Scheffel
# mcs9ff
# Module 7 HW


```{r}
library(datasets)
data(swiss)
head(swiss)
?swiss
```
# 1A

```{r}
pairs(swiss, lower.panel = NULL, main="Scatterplot of All Variables")
```

i. The Examination and Education predictors appear to be linearly related to the Fertility measure. (Perhaps an argument could be made for Agriculture.)

ii. Yes, some of the predictors appear to be highly correlated with one another:

Examinations and Education,
Examinations and Agriculture,
Agriculture and Education

# 1B

```{r}
fertilityMLR<-lm(swiss$Fertility~swiss$Agriculture+swiss$Examination+swiss$Education+swiss$Catholic+swiss$Infant.Mortality)

summary(fertilityMLR)
```

i. The ANOVA F-statistic is testing if at least one of the coefficients from the regression slopes differ from 0 (or do not), i.e. do they = 0? In context, the relevant conclusion is that the F-test (=19.76) and p-value (5.59*10^-10) < 0.05, so we reject the null hypothesis. This means the overall model is significant.

ii. The p-value does not agree with the answer to 1A in regards to the Exam variable. It does agree with the answer regarding the Education variable. The p-value and slope does not agree with the answers to part 2 of 1A (Examinations and Education, Examinations and Agriculture, Agriculture and Education). 

There is likely a strong presence of multicollinearity in the regressor variables, causing the model fit to be unreliable.

