---
title: "M03 Homework"
author: "Matt Scheffel"
date: "2022-09-19"
output:
  pdf_document: default
  html_document: default
---

# Part 1

```{r}
library(tidyverse)
copier = read.table("copier.txt", header = TRUE)
```


# 1A

The response variable in this analysis is the number of minutes spent on the phone. The predictor in this analysis is the number of copiers serviced.

# 1B

```{r}
#create scatter plot
ggplot(copier, aes(x=Minutes,y=Serviced))+
geom_point()+
labs(x="Total Minutes Spent on the Phone",
y="# of Copiers Serviced",
title="Routine Maintenance Service Calls")
```

The relationship between the 2 variables (the number of copiers serviced and the time spent by the service person) seems to be a strong positive linear relationship. The more copiers a person services, the more time they spend on the phone.

# 1C

```{r}
copierlm<-lm(Minutes~Serviced, data=copier)
summary(copierlm)
```

B^1 is = 15.0352. It is found under the Serviced Estimate.
B^0 is = -0.5802. It is found under the (Intercept) Estimate.
R^2 is = 0.9565. It is found in the section where it reads "R-squared." I used the Adjusted R-squared.
o^2 is = (8.914)^2. It is square of the residual std. error. 

# 1D

Contextual interpretation: B0 does not make sense, as a negative value would indicate that a person has negative amount of calls. The minimum should be 0.

# 1E

```{r}
anova.copier<-anova(copierlm)
anova.copier
```

F-statistic: 968.66

Null and alternative hypotheses: H0 : B1 = 0; Ha : B1 /= 0

Relevant conclusion: The p-value is less than 0.05 so we reject the null hypothesis. The data supports the hypothesis that there is a linear relationship between number of machines serviced and minutes spent on the phone.

# Part 2

# 2A

See picture attached.

# 2B

See picture attached.

# 2C

Variance is = 38

# 2D

R^2 = 0.9733

# 2E

F0 = 145.6

Relevant conclusion: The p-value is less than 0.05 so we reject the null hypothesis.

# Part 3

See picture attached.
