---
title: "HW 09"
author: "Matt Scheffel"
date: "2022-10-31"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(MASS)
data(birthwt)
head(birthwt)
```

# 1A

```{r}
df <- MASS::birthwt
df$race <- as.factor(df$race)
df$race <- as.factor(birthwt$race)

library(ggplot2)

ggplot(df, aes(x=age,y=bwt, color=race))+
geom_point()+
#scale_size(range = c(0.1,12))+
geom_smooth(method=lm, fill=NA)
labs(x="Age",
y="Birth Weight",
title="Birthweight Against Age",
addRegLine=TRUE, regLineColor="blue")
```

Interaction effect:

The regression lines for age and birth weight are not parallel and they intersect with each other. This visual shows us that there is a significant interaction effect between age and race.

# 1B

```{r}
MLRbwt <- lm(bwt~age*race, data = df)
summary(MLRbwt)
```

Regression equation:

bwt = 2583.54 + 21.37(age) + 1022.79(race2) + 326.05(race3) - 62.54(age:race2) - 26.03(age:race3)

Race 1 is the reference level. For a one unit increase in age, the weight increases by 21.37 units.

Race 2 (vs. Race 1) shows that for a one unit increase in age, the weight decreases by 41.17 units.

Race 3 (vs. Race 1) shows that for a one unit increase in age, the weight decreases by 4.66 units.
