---
title: "HW Module 5"
author: "Matt Scheffel"
date: "2022-09-30"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidyverse)
library(MASS)
library(faraway)

Data<-cornnit

```

# 1A

The response variable is the Corn Yield.

The predictor is the Nitrogen Fertilizer.

```{r}
ggplot(Data, aes(x=yield,y=nitrogen))+
geom_point()+
geom_smooth(method = "lm", se=FALSE)+
labs(x="Corn Yield (Bushels per Acre)", y="Nitrogen (Pounds per Acre)",
title="Corn Yield and Nitrogen Fertilizer")
```
The scatterplot seems to indicate that there may be a linear relationship, but there are still some significant outliers. Assumptins 1 and 2 may not be met.

# 1 B

```{r}
result<-lm(yield~nitrogen, data=Data)
Data$yhat<-result$fitted.values
Data$res<-result$residuals
ggplot(Data, aes(x=yhat,y=res))+
geom_point()+
geom_hline(yintercept=0, color="red")+
labs(x="Fitted y", y="Residuals", title="Residual Plot")
```
I will consider transforming the predictor first since I am unsure about the linear relationship of the variables.


# 1C

```{r}
bc = boxcox(result, lambda=seq(0,5))

best.lam = bc$x[which(bc$y==max(bc$y))]
```

The BoxCox plot helps me determine that I will need to transform the response variable. Variance is likely not constant. 

# 1D

```{r}
##log transform y
Data$log.yield<-log(Data$yield)

##scatter plot of y* against x
ggplot(Data, aes(x=nitrogen,y=log.yield))+
geom_point()+
geom_smooth(method = "lm", se=FALSE)+
labs(x="Nitrogen Log (Pounds per Acre)", y="Corn Yield (Bushels per Acre)",
title="Corn Yield and Nitrogen Fertilizer")
```

Relationship is more linear but there are still many outliers.

```{r}
result2<-lm(Data$log.yield~Data$nitrogen, data=Data)
##create residual plot
Data$yhat2<-result2$fitted.values
Data$res2<-result2$residuals
ggplot(Data, aes(x=yhat2,y=res2))+
geom_point()+
geom_hline(yintercept=0, color="red")+
labs(x="Fitted y", y="Residuals", title="Residual Plot (with y*)")
```
Residual plot is still fairly unchanged, suggesting the need for another transformation of the predictor variable.


```{r}
##transform x
Data$log.nitrogen<-log(Data$nitrogen)
##scatterplot of y* against x*
ggplot(Data, aes(x=log.nitrogen,y=log.yield))+
geom_point()+
geom_smooth(method = "lm", se=FALSE)+
labs(x="log Corn Yield", y="log Nitrogen",
title="Corn Yield (log) and Nitrogen Fertilizer (log)")
```

This latest transformation displays a strong linear relationship and indicates we liely need no more transformations.


```{r}
summary(result2)
```

Regression equation:

y = 4.6388 + 0.0016x

```{r}
#trying to remove NA values

#Data[is.na(Data) | Data=="Inf"] = NA

#result3<-lm(Data$log.yield~Data$log.nitrogen, data=Data)
##create residual plot
#yhat3<-result3$fitted.values
#res3<-result3$residuals
#ggplot(Data, aes(x=yhat3,y=res3))+
#geom_point()+
#geom_hline(yintercept=0, color="red")+
#labs(x="Fitted y", y="Residuals", title="Residual Plot (with y* and x*)")
```




