---
title: "Mod 09 HW"
author: "Matt Scheffel"
date: "2022-11-07"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(MASS)
data(birthwt)
head(birthwt)
?birthwt
```
# 1A

The categorical variables in this data set are "low", "race", "smoke", "ht," & "ui".

```{r}
birthwt$low = factor(birthwt$low)
birthwt$race = factor(ifelse(birthwt$race == 1, "white", 
                             ifelse(birthwt$race == 2, "black", "other")))
birthwt$smoke = factor(birthwt$smoke)
birthwt$ht = factor(birthwt$ht)
birthwt$ui = factor(birthwt$ui)
```


# 1B

Yes, "low" should be dropped as a predictor. Since we are trying to predict the birth weight of the babies, having a variable which already tells us that the baby's birth weight is low does not make sense.

# 1C
```{r}
#install.packages("leaps")
library(leaps)
allreg <- regsubsets(bwt ~., data=birthwt, nbest=2)
summary(allreg)

```

# i. 

```{r}
coef(allreg, which.max(summary(allreg)$adjr2))
```

Predictors: low1, age, lwt, racewhite

# ii.

```{r}
coef(allreg, which.min(summary(allreg)$cp))
```

Predictors: low1, racewhite, smoke1, ui1

# iii.

```{r}
coef(allreg, which.min(summary(allreg)$bic))
```

Predictors: low1, ui1

For all of these, we could have dropped the "low" predictor as discussed earlier and it would not be counted for each model. The following code could be used:

birthwt = subset(birthwt, select = -c(low))

# 1D

```{r}
regnull <- lm(bwt~1, data=birthwt)

regfull <- lm(bwt~., data=birthwt)
step(regnull, scope=list(lower=regnull, upper=regfull), direction="forward")

step(regfull, scope=list(lower=regnull, upper=regfull), direction="backward")
```

The regression equation selected using backwards selection is the last one, with the lowest AIC of 2299.75:

Step:  AIC=2299.75
bwt ~ low + race + smoke + ui

# 4

PRESS Statistic:

```{r}
press <- function(regmodel) {
sum((regmodel$residuals) / (1-lm.influence(regmodel)$hat))**2
}
```

