---
title: "STAT6021 Linear Models for Data Science - Project 01"
author: 
 - Tyler Hinnendael (dcc7qe@virginia.edu) 
 - Shrikant Mishra (stm5ne@virginia.edu)
 - Matt Scheffel (mcs9ff@virginia.edu)
 - Rex Minnis (rlm4cr@virginia.edu)
date: "2022-10-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(corrplot)
```

## Section 1
## Executive Summary

### Load diamonds4.csv dataset

```{r}
Data <- read.csv("../data/diamonds4.csv", header=TRUE)
head(Data)
```

## Section 2
##  Description of the data and the variables


## Question 1 - Data Visualizations

### 1 Use data visualizations to explore how price is related to the other 
### variables (carat, clarity, color, cut), as well as how the other variables 
### may relate to each other. Address the various claims on the 
### [diamond education page on Blue Nile](https://www.bluenile.com/education/diamonds).

```{r}
ggplot(Data, aes (carat))+
  geom_bar()+
  scale_x_binned()

```

```{r}
ggplot(Data, aes (clarity))+
  geom_bar()
```

```{r}
ggplot(Data, aes (color))+
  geom_bar()
```

```{r}
ggplot(Data, aes (cut))+
  geom_bar()
```

```{r}
ggplot(Data, aes(x = carat, y = price))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(y="Price", 
       x="Carat", 
       title="Scatterplot of Diamond Price Against Carat")
```

```{r}
ggplot(Data, aes(x = clarity, y = price))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(y="Price", 
       x="Carat", 
       title="Scatterplot of Diamond Price Against Cut")
```

```{r}
ggplot(Data, aes(x = color, y = price))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(y="Price", 
       x="Carat", 
       title="Scatterplot of Diamond Price Against Cut")
```

```{r}
ggplot(Data, aes(x = cut, y = price))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(y="Price", 
       x="Carat", 
       title="Scatterplot of Diamond Price Against Cut")
```


```{r}
ggplot(Data, aes(x=carat, y=price))+
     geom_boxplot(fill="Blue")+
     labs(x="Cut", y="Price",
      title="Distribution of Diamond Prices by Carat")
```


```{r}
ggplot(Data, aes(x=clarity, y=price))+
     geom_boxplot(fill="Blue")+
     labs(x="Cut", y="Price",
      title="Distribution of Diamond Prices by Clarity")
```


```{r}
ggplot(Data, aes(x=color, y=price_logtf))+
     geom_boxplot(fill="Blue")+
     labs(x="Color", y="Price",
      title="Distribution of Diamond Prices by Color")
```



```{r}
ggplot(Data, aes(x=cut, y=price_logtf))+
     geom_boxplot(fill="Blue")+
     labs(x="Cut", y="Price",
      title="Distribution of Diamond Prices by Cut")
```

```{r}
#Data[c("carat", "price")]
num_cols_Data <- Data %>%
                   dplyr::select(carat, price)
                  

cor(num_cols_Data)

```




```{r}
ggplot(Data, aes(x=carat, y=price))+
   geom_violin()
```


```{r}
ggplot(Data, aes(x=clarity, y=price_logtf))+
   geom_violin()
```


```{r}
#ggplot(Data, aes(x=color, y=price))+
#   geom_violin()
   #+
   #stat_summary(fun.data = "mean_cl_boot", geom = "crossbar",
   #             colour = "red", width = 0.2)
```

```{r}
ggplot(Data, aes(x=cut, y=price))+
   geom_violin()
```


```{r}
Data$price_logtf <- log(Data$price)
```


```{r}

```

## Section  3

### A description of how you fitted the regression of price against carat, and 
### the conclusions reached.

### Question 2. Fit an appropriate simple linear regression for price against carat.








