---
title: "STAT6021 Linear Models for Data Science - Project 01"
author: 
 - Tyler Hinnendael (dcc7qe@virginia.edu) 
 - Shrikant Mishra (stm5ne@virginia.edu)
 - Matt Scheffel (mcs9ff@virginia.edu)
 - Rex Minnis (rlm4cr@virginia.edu)
date: "2022-10-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(corrplot)
library(patchwork)
library(reshape)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
```

## Section 1
## Executive Summary

### Load diamonds4.csv dataset

```{r}
Data <- read.csv("../data/diamonds4.csv", header=TRUE)
head(Data)
```

## Section 2
##  Description of the data and the variables


## Question 1 - Data Visualizations

### 1 Use data visualizations to explore how price is related to the other 
### variables (carat, clarity, color, cut), as well as how the other variables 
### may relate to each other. Address the various claims on the 
### [diamond education page on Blue Nile](https://www.bluenile.com/education/diamonds).

```{r}
#ggplot(Data, aes (carat))+
#  geom_bar()+
#  scale_x_binned()
```

```{r}
#skimr::skim(Data)
```

```{r}
#kbl(skimr::skim(Data)) %>%
#  kable_styling(bootstrap_options = c("striped", "hover"))
```


```{r}
#ggplot(Data, aes (clarity))+
#  geom_bar()
```

```{r}
#ggplot(Data, aes (color))+
#  geom_bar()
```

```{r}
#ggplot(Data, aes (cut))+
#  geom_bar()
```

```{r}
ggplot(Data, aes(x = carat, y = price))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(y="Price", 
       x="Carat", 
       title="Scatterplot of Diamond Price Against Carat")
```

```{r}
#ggplot(Data, aes(x = clarity, y = price))+
#  geom_point()+
#  geom_smooth(method = "lm", se=FALSE)+
#  labs(y="Price", 
#       x="Carat", 
#       title="Scatterplot of Diamond Price Against Cut")
```

```{r}
#ggplot(Data, aes(x = color, y = price))+
#  geom_point()+
#  geom_smooth(method = "lm", se=FALSE)+
#  labs(y="Price", 
#       x="Carat", 
#       title="Scatterplot of Diamond Price Against Cut")
```

```{r}
#ggplot(Data, aes(x = cut, y = price))+
#  geom_point()+
#  geom_smooth(method = "lm", se=FALSE)+
#  labs(y="Price", 
#       x="Carat", 
#       title="Scatterplot of Diamond Price Against Cut")
```


```{r}
ggplot(Data, aes(x=carat, y=price))+
     geom_boxplot(fill="Blue")+
     labs(x="Cut", y="Price",
      title="Distribution of Diamond Prices by Carat")
```


```{r}
ggplot(Data, aes(x=clarity, y=price))+
     geom_boxplot(fill="Blue")+
     labs(x="Cut", y="Price",
      title="Distribution of Diamond Prices by Clarity")
```
```{r}
Data$price_logtf <- log(Data$price)
```

```{r}
ggplot(Data, aes(x=color, y=price_logtf))+
     geom_boxplot(fill="Blue")+
     labs(x="Color", y="Price",
      title="Distribution of Diamond Prices by Color")
```

# Question 2 - Univariate Analysis: Frequency Plots

```{r}
#create the log(price) column to transform the y variable
vec <- log(Data$price)
Data["log_price"] <- vec

# reorders these two columns in the diamonds dataframe so it plots in descending order.
Data$clarity <- factor(Data$clarity, levels = c("FL", "IF", "VVS1", "VVS2","VS1","VS2", "SI1"))
Data$cut <- factor(Data$cut, levels = c("Astor Ideal", "Ideal", "Very Good", "Good"))

p1 <- ggplot(Data, aes(carat))+
  geom_histogram(binwidth = 0.5, fill = "#0073C2FF")+
  ggtitle("Frequency of Diamond Weights in Carats")+
  xlab("Diamond Weight (carat)")+ 
  ylab("Count")+ 
  theme(plot.title = element_text(size=11))

p2 <- ggplot(Data, aes(cut))+
  geom_bar(fill = "#0073C2FF")+
  ggtitle("Frequency of Diamond Cut Grades")+
  xlab("Cut (grades)")+ 
  ylab("Count")+ 
  theme(plot.title = element_text(size=11))

p3 <- ggplot(Data, aes(color))+
  geom_bar(fill = "#0073C2FF")+
  ggtitle("Frequency of Diamond Color Grades")+
  xlab("Color (grades)")+ 
  ylab("Count")+ 
  theme(plot.title = element_text(size=11))

p4 <- ggplot(Data, aes(clarity))+
  geom_bar(fill = "#0073C2FF")+
  ggtitle("Frequency of Diamond Clarity Grades")+
  xlab("Clarity (grades)")+ 
  ylab("Count")+ 
  theme(plot.title = element_text(size=11))

p1 + p2 + p3 + p4 + plot_layout(ncol = 2)
```



```{r}
cut_mean <- aggregate(Data$log_price, list(Data$cut), FUN=mean)
cols <- c("Cut", "Mean_log_Price")
colnames(cut_mean) <- cols
cut_mean$Cut <- factor(cut_mean$Cut, levels=c("Astor Ideal", "Ideal", "Very Good", "Good"))

color_mean <- aggregate(Data$log_price, list(Data$color), FUN=mean)
cols <- c("Color", "Mean_log_Price")
colnames(color_mean) <- cols


clarity_mean <- aggregate(Data$log_price, list(Data$clarity), FUN=mean)
cols <- c("Clarity", "Mean_log_Price")
colnames(clarity_mean) <- cols
clarity_mean$Clarity<- factor(clarity_mean$Clarity, levels=c("FL", "IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2"))


g1 <- ggplot(Data, aes(x = carat, y = price))+
        geom_point()+
        geom_smooth(method = "lm", se=FALSE)+
        labs(y="Price ($)", x="Weight of the diamond (carats)", 
           title="Scatterplot of Diamond Price Against Carat")

g2 <- ggplot(cut_mean, aes(x=Cut, y=Mean_log_Price))+
  geom_bar(stat = "identity", fill = "coral", alpha = 0.5)+
  ggtitle("Mean Log(Price) for Diamond Cut Grades")+
  xlab("Cut (grades)")+ 
  ylab("Mean Log Price ($)")+ 
  theme(plot.title = element_text(size=11))

g3 <- ggplot(color_mean, aes(x=Color, y=Mean_log_Price))+
  geom_bar(stat = "identity", fill = "coral", alpha = 0.5)+
  ggtitle("Mean Log(Price) for Diamond Color Grades")+
  xlab("Color (grades)")+ 
  ylab("Mean Log Price ($)")+ 
  theme(plot.title = element_text(size=11))

g4 <- ggplot(clarity_mean, aes(x=Clarity, y=Mean_log_Price))+
  geom_bar(stat = "identity", fill = "coral", alpha = 0.5)+
  ggtitle("Mean Log(Price) for Diamond Clarity Grades")+
  xlab("Clarity (grades)")+ 
  ylab("Mean Log Price ($)")+ 
  theme(plot.title = element_text(size=11))

g1 + g2 + g3 + g4 + plot_layout(ncol = 2)

```



```{r}

#ggplot(Data, aes(x=cut, y=log_price, fill=cut)) +
#  geom_violin() +
#  geom_jitter(width=0, alpha=0.5) +
#  scale_fill_brewer(palette="Blues", direction=-1)

ggplot(Data, aes(x=color, y=log_price, fill=color)) +
  geom_violin() +
  geom_jitter(width=0, alpha=0.5) +
  scale_fill_brewer(palette="Blues", direction=-1)+
  labs(y="Log(Price) ($)", x="Color (grade)", 
           title="Log(Price) for Each Grade of Diamond Color")

#ggplot(Data, aes(x=clarity, y=log_price, fill=clarity)) +
#  geom_violin() +
#  geom_jitter(width=0, alpha=0.5) +
#  scale_fill_brewer(palette="Blues", direction=-1)
```



```{r}

# created 'price per carat" to normalize based on a wide range of carat weights
vec <- (Data$price / Data$carat)
Data["price_per_carat"] <- vec

# A new dataframe 'varmeans' is created to represent the 'mean price_per_carat' for all cut/color/clarity combinations of the diamonds.
# This dataframe is used for the multivariate plots.
var_means <- aggregate(Data$price_per_carat, list(Data$cut, Data$color, Data$clarity), FUN=mean)
cols <- c("Cut", "Color", "Clarity", "Mean_Price_Per_Carat")
colnames(var_means) <- cols
var_means$Clarity<- factor(var_means$Clarity, levels=c("FL", "IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2"))
var_means$Cut<- factor(var_means$Cut, levels=c("Astor Ideal", "Ideal", "Very Good", "Good"))
```

```{r}
# Outputs comparisons of Color and Cut combinations
p<-ggplot(var_means, aes(x=Color, y=Mean_Price_Per_Carat, fill=Cut)) +
  geom_bar(position="dodge", stat = "identity") +
  scale_fill_brewer(palette = 7, direction = -1) +
  scale_colour_brewer(palette = 1)
p+theme_dark() + ggtitle("Mean Price-per-Carat for Grade Combinations of Diamond Color and Cut") +
  xlab("Color (grade)") + ylab("Mean Price-per-Carat ($)")
```


```{r}
# Outputs comparisons of Color and Clarity combinations
p<-ggplot(var_means, aes(x=Color, y=Mean_Price_Per_Carat, fill=Clarity)) +
  geom_bar(position="dodge", stat = "identity") +
  scale_fill_brewer(palette = 7, direction = -1) +
  scale_colour_brewer(palette = 1)
p+theme_dark() + ggtitle("Mean Price-per-Carat for Grade Combinations of Diamond Color and Clarity") +
  xlab("Color (grade)") + ylab("Mean Price-per-Carat ($)")
```

## Section  3

### A description of how you fitted the regression of price against carat, and 
### the conclusions reached.

### Question 2. Fit an appropriate simple linear regression for price against carat.


```{r}
ggplot(Data, aes(x = carat, y = price))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
    labs(y="Price", 
       x="Carat", 
       title="Scatterplot of Diamond Price Against Carat")
```



```{r}
results <- lm(price~carat, data=Data)
summary(results)
```


```{r}
yhat <- results$fitted.values
res <- results$residuals

ggplot(Data, aes(x=yhat, y=res))+
  geom_point()+
  geom_hline(yintercept = 0, color="red")+
  labs(x="Fitted y", y="Residuals", title="Residual Plot with yhat")
```

```{r}
boxcox(results, lambda = seq(0, 1, 0.1))
```


```{r}
ystar <- log(Data$price)
results2 <- lm(ystar~carat, data=Data)
yhat2 <- results2$fitted.values
res2 <- results2$residuals

ggplot(Data, aes(x=yhat2, y=res2))+
  geom_point()+
  geom_hline(yintercept = 0, color="red")+
  labs(x="Fitted y", y="Residuals", title="Residual Plot with ystar")
```

```{r}
ggplot(Data, aes(x=carat, y=ystar))+
  geom_point()+
  geom_smooth(method="lm", se=FALSE)+
  labs(y="ystar log(Price)", 
       x="Carat", 
       title="Scatterplot of Diamond ystar Against Carat")
```

```{r}
xstar <- log(Data$carat)
results3 <- lm(ystar~xstar, data=Data)
yhat3 <- results3$fitted.values
res3 <- results3$residuals

ggplot(Data, aes(x=yhat3, y=res3))+
  geom_point()+
  geom_hline(yintercept = 0, color="red")+
  labs(x="Fitted y", y="Residuals", title="Residual Plot with ystar and xstar")

```


```{r}
summary(results3)
```


```{r}
cor(xstar, ystar)
```


```{r}
qqnorm(res3)
qqline(res3, col = "red")
```



```{r}
a <- acf(res3, main="ACF Plot of Residuals with ystar and xstar")

```

```{r}
ggplot(Data, aes(x=xstar, y=ystar))+
  geom_point()+
  geom_smooth(method="lm", se=FALSE)+
  labs(y="ystar log(Price)", 
       x="xstar log(Carat)", 
       title="Scatterplot of Diamond ystar Against xstar")
```

