---
title: "Project 1"
author: "Tyler Hinnendael"
date: "2022-10-11"
output: html_document
---

# Question 1

-- to be filled out in the high-level report.

# Question 2
Prompt:
a) A description of the data and the variables
  -- created a variable "price per carat", to combine these two features so the visualizations 
  take into account the 'weight' of the diamond size in the price.
b) the data visualizations you created to address how price is related to the other variables
  -- violin plot to show distribution of mean-price against each feature
  -- bar chart with mean price trends for each category/sub-category
  -- bar plot to show mean-price trends for each combination of the 3 categorical variables
  -- could not do a correlation matrix because 3 of the variables are not numeric
c) the claims made on the diamond education page. 
  -- " Cut is considered the most important of the four C's. Color is the second most important of the 4Cs." 
  -> Color appears to be the most consistent feature across all observations. Cut and clarity have some influential
  observations and categories (flawless clarity). Cut does not appear to be the most important feature.
d) Provide contextual commentary on the visualizations.

```{r echo=FALSE}
## reads in the diamonds4.csv file
diamonds<-read.csv("diamonds4.csv", header=TRUE)
```
```{r echo=FALSE}
# created 'price per carat" to normalize based on a wide range of carat weights
vec<- (diamonds$price / diamonds$carat)
diamonds["price_per_carat"]<-vec
```
```{r echo=FALSE}
# created the log(price) column to transform the y-variable. 
vec<- log(diamonds$price)
diamonds["log_price"]<-vec
```

# (Question 2) - Univariate Analysis: Frequency Plots

```{r echo=FALSE}
# reorders these two columns in the diamonds data frame so it plots in descending order.
diamonds$clarity<- factor(diamonds$clarity, levels=c("FL", "IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2"))
diamonds$cut<- factor(diamonds$cut, levels=c("Astor Ideal", "Ideal", "Very Good", "Good"))
```

```{r}
print("min price: ")
min(diamonds$price)
print("max price: ")
max(diamonds$price)
print("mean price")
mean(diamonds$price)
median(diamonds$price)

print("min carat weight: ")
min(diamonds$carat)
print("max carat weight: ")
max(diamonds$carat)
print("mean weight")
mean(diamonds$carat)
median(diamonds$carat)
```


```{r echo=FALSE}
# Creates univariate plots for each of the 4 C's by counting the number of observations for each sub-category (grade)
library(ggplot2)
p1<-ggplot(diamonds, aes(carat)) + 
  geom_histogram(binwidth = 0.5, fill = "#0073C2FF")+ 
  ggtitle("Frequency of Diamond Weights in Carats")+
  xlab("Diamond Weight (carat)") + ylab("Count")+ theme(plot.title = element_text(size=11))

p2<-ggplot(diamonds, aes(cut)) + 
  geom_bar(fill = "#0073C2FF") + 
  ggtitle("Frequency of Diamond Cut Grades")+
  xlab("Cut (grades)") + ylab("Count")+ theme(plot.title = element_text(size=11))

p3<-ggplot(diamonds, aes(color)) + 
  geom_bar(fill = "#0073C2FF") + 
  ggtitle("Frequency of Diamond Color Grades")+
  xlab("Color (grades)") + ylab("Count")+ theme(plot.title = element_text(size=11))

p4<-ggplot(diamonds, aes(clarity)) + 
  geom_bar(fill = "#0073C2FF") + 
  ggtitle("Frequency of Diamond Clarity Grades")+
  xlab("Clarity (grades)") + ylab("Count")+ theme(plot.title = element_text(size=11))
```
```{r echo=FALSE}
# Outputs the Univariate Plots
library(patchwork)
p1 + p2 + p3 + p4 + plot_layout(ncol = 2)
```

# (Question 2) - Univariate Analysis: Frequency Tables 

```{r echo=FALSE}
# Creates frequency tables for Cut, Color, and Clarity
freq_cut<-table(diamonds$cut)
freq_color<-table(diamonds$color)
freq_clarity<-table(diamonds$clarity)

# Prints the Cut frequency table
order_ind<-c("Astor Ideal", "Ideal", "Very Good", "Good")
print("Frequency Table for Grades of Diamond Cut")
freq_cut[order_ind]
```
```{r echo=FALSE}
# Prints the Color frequency table
print("Frequency Table for Grades of Diamond Color")
print(freq_color)
```
```{r echo=FALSE}
# Prints the Clarity frequency table
order_ind<-c("FL", "IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2")
print("Frequency Table for Grades of Diamond Clarity")
freq_clarity[order_ind]
```


# (Question 2) - Bivariate Analysis: Bar charts of mean(log(price)) for each cut, color, and clarity grade. Scatterplot of Price and Carat sizes.

```{r echo=FALSE}
library(ggplot2)

# This code block creates a dataframe for cut, color and clarity. The mean(log(price)) is calculated for each grade of cut, color and clarity.
# These data frames are used for the bivariate plots of Price against each of the C's. 
cut_mean<-aggregate(diamonds$log_price, list(diamonds$cut), FUN=mean)
cols <- c("Cut", "Mean_log_Price")
colnames(cut_mean) <- cols
cut_mean$Cut<- factor(cut_mean$Cut, levels=c("Astor Ideal", "Ideal", "Very Good", "Good"))

color_mean<-aggregate(diamonds$log_price, list(diamonds$color), FUN=mean)
cols <- c("Color", "Mean_log_Price")
colnames(color_mean) <- cols

clarity_mean<-aggregate(diamonds$log_price, list(diamonds$clarity), FUN=mean)
cols <- c("Clarity", "Mean_log_Price")
colnames(clarity_mean) <- cols
clarity_mean$Clarity<- factor(clarity_mean$Clarity, levels=c("FL", "IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2"))
```

```{r echo=FALSE}
# Creates a bivariate plot for each of the 4 C's
g1<-ggplot(diamonds, aes(x=carat, y=price))+
  geom_point()+
  labs(x="Weight of the diamond (carats)", y="price ($)",
       title="Scatterplot of Price against Carats") + theme(plot.title = element_text(size=11))

g2<-ggplot(cut_mean, aes(x=Cut, y=Mean_log_Price)) + geom_bar(stat = "identity", fill = "coral", alpha = 0.5) + 
  ggtitle("Mean log(Price) for Diamond Cut Grades") +
  xlab("Cut (grade)") + ylab("Mean Log Price ($)") + theme(plot.title = element_text(size=11))

g3<-ggplot(color_mean, aes(x=Color, y=Mean_log_Price)) + geom_bar(stat = "identity", fill = "coral", alpha = 0.5) + 
  ggtitle("Mean log(Price) for Diamond Color Grades") +
  xlab("Color (grade)") + ylab("Mean Log Price ($)")+ theme(plot.title = element_text(size=11))

g4<-ggplot(clarity_mean, aes(x=Clarity, y=Mean_log_Price)) + geom_bar(stat = "identity", fill = "coral", alpha = 0.5) + 
  ggtitle("Mean log(Price) for Diamond Clarity Grades") +
  xlab("Clarity (grade)") + ylab("Mean Log Price ($)")+ theme(plot.title = element_text(size=11))
```

```{r echo=FALSE}
# Outputs the four bivariate plots.
library(patchwork)
g1 + g2 + g3 + g4 + plot_layout(ncol = 2)
```

```{r echo=FALSE}
# Outputs the distribution of log(Price) values for each Diamond Color Grade
ggplot(diamonds, aes(x=color, y=log_price, fill=color)) +
  geom_violin() +
  geom_jitter(width=0.1, alpha=0.5) +
  scale_fill_brewer(palette="Blues", direction=-1)+ 
  ggtitle("Log(Price) Distribution for Each Grade of Diamond Color")+
  xlab("Color (grade)") + ylab("Log(Price) ($)")
```

# (Question 2) - Multivariate Plots: ggplots showing trends in mean-price per carat with combinations of the 3 categorical C's

```{r echo=FALSE}
library(MASS)
library(reshape)
library(reshape2)

# A new dataframe 'varmeans' is created to represent the 'mean price_per_carat' for all cut/color/clarity combinations of the diamonds.
# This dataframe is used for the multivariate plots.
var_means<-aggregate(diamonds$price_per_carat, list(diamonds$cut, diamonds$color, diamonds$clarity), FUN=mean)
cols <- c("Cut", "Color", "Clarity", "Mean_Price_Per_Carat")
colnames(var_means) <- cols
var_means$Clarity<- factor(var_means$Clarity, levels=c("FL", "IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2"))
var_means$Cut<- factor(var_means$Cut, levels=c("Astor Ideal", "Ideal", "Very Good", "Good"))
```

```{r echo=FALSE}
# Outputs comparisons of Color and Cut combinations
p<-ggplot(var_means, aes(x=Color, y=Mean_Price_Per_Carat, fill=Cut)) +
  geom_bar(position="dodge", stat = "identity") +
  scale_fill_brewer(palette = 7, direction = -1) +
  scale_colour_brewer(palette = 1)
p+theme_dark() + ggtitle("Mean Price-per-Carat for Grade Combinations of Diamond Color and Cut") +
  xlab("Color (grade)") + ylab("Mean Price-per-Carat ($)")
```
```{r echo=FALSE}
# Outputs comparisons of Color and Clarity combinations
p<-ggplot(var_means, aes(x=Color, y=Mean_Price_Per_Carat, fill=Clarity)) +
  geom_bar(position="dodge", stat = "identity") +
  scale_fill_brewer(palette = 7, direction = -1) +
  scale_colour_brewer(palette = 1)
p+theme_dark() + ggtitle("Mean Price-per-Carat for Grade Combinations of Diamond Color and Clarity") +
  xlab("Color (grade)") + ylab("Mean Price-per-Carat ($)")
```

